Networks Matter in International Relations:\\A Latent Network Approach to Dyadic Analysis

### 2017-10-24 paper/5_concl.tex
small stuff.
small changes throughout

\section{Conclusion}

International relations is generally about the interactions and dependences among a set of countries or other important actors such as international governmental organizations (IGOs, such as the WTO) and non-governmental organizations (NGOs, such as the Gates Foundation).  Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time. This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citet{erikson:pinto:2014,aronow:etal:2015}.

@@

To explore this approach in the context of international relations we have presented two broad analyses. The first is a simulation of where the characteristics of the network are known. This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models). The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.  These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.  We then reanalyzed each study using the latent network approach which captures that additive and multiplicative aspects of interdependencies among the dyadic data.  In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.  This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they only obtain under stringent assumptions that can not possibly be valid.  This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.

>>>>

\section{\textbf{Conclusion}}

International relations is generally about the interactions and dependences among a set of countries or other important actors such as international governmental organizations (IGOs, such as the WTO) and non-governmental organizations (NGOs, such as the Gates Foundation).  Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time. This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.

@@

To explore this approach in the context of international relations we have presented two broad analyses. The first is a simulation where the characteristics of the network are known. This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models). The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.  These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.  We then reanalyzed each study using the latent network approach which captures additive and multiplicative aspects of interdependencies among the dyadic data.  In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.  This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that can not possibly be valid.  This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.

### 2017-10-24 paper/netsMatter.tex
starting to rework the app section.
stargazer tables
more clean up we are at 99%
first cut of sim done
appendix touchups
share for comments
more tiny changes
the coeff tables need to have the st. errors put in math mode $$; OR, we need to take the estimated coefficients out of math mode.
small changes throughout

% Fonts
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage[T1]{fontenc}
\usepackage{ae}

@@

% % add watermark
% \usepackage{draftwatermark}
% \SetWatermarkText{Draft: Not Ready}
% \SetWatermarkScale{.5}
% \SetWatermarkColor[gray]{0.9}

@@

@@

\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
\vspace{\baselineskip}

@@

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
\chead{\fancyplain{}{How to fix IR}}
\rhead{\fancyplain{}{\today}}
\rfoot{\fancyplain{}{\thepage}}
% references to graphics
\makeatletter
% let's keep them in the same directory \def\input@path{{}

@@

\small{\singlespacing{
\begin{abstract}
	International relations scholarship is filled with dyads. This includes dyadic hypotheses and propositions, but especially data. These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads. However, most empirical studies of such data do not take into account these dependencies. As a result, the studies are often contradictory and produce results that are not compelling. One reason is that the independence required by the typical methods employed does not exist in the data being studied.  We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies. We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach. The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
\end{abstract}
}}

@@

\newpage
% Bib stuff
\clearpage
\bibliography{/Users/s7m/whistle/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}

>>>>

% Fonts
% at some point figure out bolding ...
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage[T1]{fontenc}
\usepackage{ae}

@@

% add watermark
\usepackage{draftwatermark}
\SetWatermarkText{Version 1.0}
\SetWatermarkScale{.5}
\SetWatermarkColor[gray]{0.9}

@@

% outside code
\usepackage{listings}
\usepackage{enumitem}

@@

\title{Networks Matter in International Relations:\\A Latent Network Approach to Dyadic Analysis}
\vspace{\baselineskip}

@@

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
\chead{\fancyplain{}{Networks Matter in IR}}
\rhead{\fancyplain{}{\today}}
\rfoot{\fancyplain{}{\thepage}}
% references to graphics
\makeatletter
% let's keep them in the same directory \def\input@path{{}

@@

\small{\singlespacing{
\begin{abstract}
	International relations scholarship is filled with dyads. This includes dyadic hypotheses and propositions, but especially data. These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads. However, most empirical studies of such data do not take into account these dependencies. As a result, the studies are often contradictory and produce findings that are not compelling. One reason is that the independence required by typical methods does not exist in the data. We present a different, regression based method, which constructs a latent network and incorporates \first, \second, and \third-order dependencies. We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach. The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects and it also dominates standard approaches in terms of out-of-sample cross-validations.
\end{abstract}
}}

@@

\newpage
% Bib stuff
\clearpage
%\bibliography{/Users/s7m/whistle/master}
\bibliography{/Users/mdw/Git/whistle/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}
